---
title: "Exploratory analyses on procrastination data using cluster analysis"
output: 
  html_document:
    toc: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```

<br>
<br>

# Cluster analysis

<br>

### Import data
```{r import_dat}
load("totscores_modif.Rda")
```


```{r import_lib}
library(cluster)
library(factoextra)
library(GGally)
library(tidyr)
library(plotly)
library(NbClust)
library(clValid)
library(dendextend)
library(purrr)
library(psych)
library(car)
library(fpc)
```
<br>

### Preparing data for cluster analysis
```{r data_prep}
## Load data
load("totscores_modif.Rda") 

### Outlier analysis

out_vec <- totscores_mod[ ,c(24:32)]

mahal <- mahalanobis(
    out_vec,
    colMeans(out_vec, na.rm = T),
    cov(out_vec, use = "pairwise.complete.obs")
)


cutoff <-
    qchisq(1 - .001, ncol(out_vec))

summary(mahal < cutoff)

noout_dat <- totscores_mod %>% 
    filter(mahal < cutoff) 

# isolate clustering variables (only school functioning variable)

vars_vec <- names(totscores_mod[ ,c(24, 29:32)])

vars_vec

clust_dat <- noout_dat[ ,vars_vec]

# Standardize the variables
scaled_dat <- scale(clust_dat)

# Compute the dissimilarity matrix
res.dist <- dist(scaled_dat, method = "euclidean")
dist_matrix <- as.matrix(res.dist)
```
<br>

### Compare estimation methods
```{r compare}
m <- c( "average", "single", "complete", "ward")
names(m) <- c( "average", "single", "complete", "ward")

# function to compute coefficient
ac <- function(x) {
    agnes(scaled_dat, method = x)$ac
}

map_dbl(m, ac)
```

The analysis shows the superiority of the Ward method in this case.

<br>

### Determining the optimal number of clusters

```{r nb_clust}
res.nbclust <-
    NbClust(
        scaled_dat,
        distance = "euclidean",
        min.nc = 2,
        max.nc = 9,
        method = "complete",
        index = "all"
    )

factoextra::fviz_nbclust(res.nbclust) + 
    theme_minimal() + 
    ggtitle("NbClust's optimal number of clusters")
```


## Cluster analyses

<br>

### Hierarchical K-Means Clustering
```{r hkmeans}

res.hk <- hkmeans(dist_matrix, 2)

# Visualize the hkmeans final clusters
fviz_cluster(
    res.hk,
    cex = .6,
    palette = "jco",
    repel = T,
    ggtheme = theme_classic()
)

# Visualize the tree
fviz_dend(res.hk, cex = 0.6, palette = "jco",
          rect = TRUE, rect_border = "jco", rect_fill = TRUE)

# Create a group variable based on clustering
grp_res.hk <- res.hk$cluster
totscores_mod$cluster_res.hk <- grp_res.hk

# Save the grouping variable
save(totscores_mod, file = "totscores_w_cluster.Rda")

```
<br>
<br>

# Exploratory comparisons between clusters 

<br>

### Import libraries

```{r library}
library(tidyverse)
library(ggplot2)
```
<br>

### Import data files with clustering variable
```{r import data}
load(file = "totscores_w_cluster.Rda")
```
<br>

### Create factor variables from cluster variables and "proc_ouvert" variable
```{r data wrangling}

totscores_mod <- totscores_mod %>% 
    mutate(cluster_res.hk = factor(cluster_res.hk,
                            labels = c("1", "2"))) %>% 
    mutate(cluster_res.diana = factor(cluster_res.diana,
                                      labels = c("1", "2"))) %>% 
    mutate(proc_ouv = factor(proc_ouvert,
                             labels = c("À diminué",
                                        "Demeuré le même",
                                        "Empiré beaucoup",
                                        "Empiré quelque peu")))
```
<br>
<br>

## Visualize differences based on clustering variable

```{r plot}

totscores_mod %>% 
    ggplot(aes(proc_ouv, fill = cluster_res.hk)) +
    geom_histogram(stat = "count", binwidth = .5, alpha = .5) +
    facet_wrap(~cluster_res.hk, ncol = 1) +
    theme(axis.text.x = element_text(angle = 90)) +
    theme_bw()

totscores_mod %>% 
    ggplot(aes(proc_ouv, fill = cluster_res.diana)) +
    geom_histogram(stat = "count", binwidth = .5, alpha = .5) +
    facet_wrap(~cluster_res.diana, ncol = 1) +
    theme(axis.text.x = element_text(angle = 90)) +
    theme_bw()

```
<br>
<br>

## Exploratory analyses : mean differences
<br>

### Import library
```{r import_lib3}
library(rstatix)
```


```{r mean_sd}

### Look for differences in perfectionnisme
totscores_mod %>% 
    t_test(qpr_pp ~ cluster_res.hk)

totscores_mod %>% 
    t_test(qpr_rhs ~ cluster_res.hk)

totscores_mod %>% 
    group_by(cluster_res.hk) %>% 
    get_summary_stats(qpr_rhs)
```

Différence significative sur la variable Recherche de hauts standards avec une taille d'effet de : `totscores_mod %>% cohen_d(qpr_rhs ~ cluster_res.hk)`

<br>

```{r mean_sd_gses}

### Look for differences in self-efficacy
totscores_mod %>%
    t_test(gses_tot ~ cluster_res.hk)

totscores_mod %>% 
    group_by(cluster_res.hk) %>% 
    get_summary_stats(gses_tot)
```

Différence significative sur la variable Auto efficacité avec une taille d'effet de : `totscores_mod %>% cohen_d(gses_tot ~ cluster_res.hk)`

```{r mean_sd_mpfi}

### Look for differences in psychological flexibility
totscores_mod %>%
    t_test(mpfi_tot ~ cluster_res.hk)

totscores_mod %>% 
    group_by(cluster_res.hk) %>% 
    get_summary_stats(mpfi_tot)
```

Différence significative sur la variable Flexibilité psychologique avec une taille d'effet de : `totscores_mod %>% cohen_d(mpfi_tot ~ cluster_res.hk)`

<br>
<br>

## Exploratory analyses : mean differences based on PPS cutoffs
<br>


```{r pps_cutoffs}

### Create group variable based on cutoffs
totscores_mod <- totscores_mod %>% 
    mutate(pps_cutoffs = case_when(
        pps_tot <= 27 ~ "low", 
        (pps_tot > 27 & pps_tot <= 39) ~ "medium",
        (pps_tot > 39) ~ "high")) 

totscores_mod <- totscores_mod %>% 
    mutate(pps_cutoffs = factor(pps_cutoffs,
                                levels = c("low", "medium", "high")))

```




```{r pps_cutoffs_mean_diff}

totscores_mod %>%
    anova_test(mpfi_tot ~ pps_cutoffs)

totscores_mod %>%
    anova_test(gses_tot ~ pps_cutoffs)

totscores_mod %>%
    anova_test(qmes_eng_opt ~ pps_cutoffs)

totscores_mod %>%
    anova_test(qmes_sur_eng_scol ~ pps_cutoffs)

totscores_mod %>%
    anova_test(qmes_sous_eng_scol ~ pps_cutoffs)

totscores_mod %>%
    anova_test(ese_tot ~ pps_cutoffs)
```

